---
title: "PRSs after CABG & PCI"
editor_options:
  chunk_output_type: console
output:
  html_document:
    number_sections: true
---

```{r setup, include=FALSE}
#packrat::init()
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```


**Libraries**


```{r}
#setwd("/finngen/red/cabg_prs")

#install.packages("/home/ivm/R/src/packrat_0.5.0.tar.gz", repos=NULL, type="source")

packrat::on()
library(tidyverse)
library(survival)   # survival analysis

library(data.table) # fread() function
library(gridExtra)  # plots to grid
library(survminer)  # ggbased visualization and extra diagnostics
library(survRM2)    # restricted mean survival times
library(visdat)     # visualization of tibble and na's
library(forestplot) # forestplot
library(compareC)   # compare c-index
library(kableExtra) # Pretty tables
library(tableone)   # Characteristics table

source('/finngen/red/cabg_prs/scripts/functions2.R')

data_path <- "/finngen/red/cabg_prs/data"
prs_path <- "/finngen/red/cabg_prs/data/ukb_prs"
fig_path <- "/finngen/red/cabg_prs/figs"

```


# Data

## Variables

**Main endpoints**

* I9_CABG, Coronary artery bypass crafting
* I9_ANGIO, Coronary angiopasty
* I9_REVASC, I9_CABG OR I9_ANGIO

Three samples: patients *after* CABG, ANGIO and REVASC.

**For polygenic riskscores**

* I9_AF (ukb: 20002_1471), Atrial fibrillation
* I9_CHD,	Major coronary heart disease event
* I9_INTRACRA,	Nontraumatic intracranial haemmorrhage
* I9_MI,	Myocardial infarction
* I9_STR,	Stroke, excluding SAH
* K11_GIBLEEDING,	GI-bleeding
 
**For clinical riskscore CHADSVASc:**
* Congestive heart failure = I9_HEARTFAIL
* Hypertension = I9_HYPTENS
* Diabetes mellitus = E4_DIABETES
* Stroke, TIA, or TE = C_STROKE
* Vascular disease = I9_MI or I9_PAD

## Download & preprocess data


**Copying and preprocessing outside R**
Endpoint and covariate files transfered and unzipped. Columns selected from endpoint file prior to import to r, because the original phenotype file is very large.


```{bash, eval=F}

#Endpoint file and covariate files unzipped
phenodir="/finngen/red/cabg_prs/data/phenotypes"
gzip -dc /finngen/library-red/finngen_R6/phenotype_2.0/data/finngen_R6_v2_endpoint.gz > $phenodir/finngen_R6_v2_endpoint
gzip -dc /finngen/library-red/finngen_R6/phenotype_2.0/data/finngen_R6_cov_pheno_1.0.txt.gz > $phenodir/finngen_R6_cov_pheno_1.0.txt

#Self written perl-script to extract columns from phenotype file (runned from directory 'scripts')
phenodir="/finngen/red/cabg_prs/data/phenotypes"
perl select_columns.pl $phenodir/finngen_R6_v2_endpoint       $phenodir/finngen_R6_v2_CABG    fg_pheno_cols.txt &
perl select_columns.pl $phenodir/finngen_R6_cov_pheno_1.0.txt $phenodir/R6_cov_pheno_selected fg_cov_cols.txt &
perl select_columns.pl $phenodir/finngen_R6_v2_endpoint       $phenodir/finngen_R6_v2_chadsvasc    fg_chadsvasc_cols.txt &


```


**Loading and combining the data**

Downloading the data

```{r}

names_in <- list("I9_AF", "I9_CHD", "I9_INTRACRA", "I9_MI", "I9_STR", "K11_GIBLEEDING")

#PRS scores
prs_df <- lapply(names_in, function(name) {
  fread(str_glue("{prs_path}/ukb.{name}.both_sexes.profile")) %>%
        select(IID, SCORE) %>%
      rename(FINNGENID = IID, !!paste(name, "SCORE", sep="_") := SCORE)}
  ) %>%
  reduce(left_join, by = "FINNGENID")


#endpoints and covariates
phenotypes <- fread(str_glue("{data_path}/phenotypes/finngen_R6_v2_CABG"), sep = "\t") 
covs_in <- fread(str_glue("{data_path}/phenotypes/R6_cov_pheno_selected"), sep = "\t")
pheno_chdvsc <- fread(str_glue("{data_path}/phenotypes/finngen_R6_v2_chadsvasc"), sep = "\t") %>%
  select(-contains('MI')) #To avoid duplicates

```

Combining the data

```{r}
df <- phenotypes %>%
  left_join(prs_df, by = "FINNGENID") %>%
  left_join(covs_in, by = "FINNGENID") %>%
  left_join(pheno_chdvsc, by = "FINNGENID") %>%
  as_tibble() %>%
  mutate_at(c("batch","cohort"),as.factor) 

```


**Preprocess**

Removing individuals without genetic information
Also endpointnames are simplified.

```{r}
#
#glimpse(df)
#vis_dat(sample_n(df, 10000))
df <- df %>% filter(!is.na(batch)) %>%
  rename_at(vars(contains('I9_')), list(~str_remove(., "I9_"))) %>%
  rename_at(vars(contains('K11_')), list(~str_remove(., "K11_"))) %>%
  rename_at(vars(contains('E4_')), list(~str_remove(., "E4_"))) %>%
  rename_at(vars(contains('C_STRO')), list(~str_remove(., "C_"))) %>%
  rename_at(vars(contains('ANGIO')), list(~str_replace(., "ANGIO", "PCI")))
#vis_dat(sample_n(df, 10000))

#Number of individuals:
#dim(df)[1]

```


Creating scaled & categorical variables


```{r}

#my_probs = c(0 ,0.2, 0.8, 1)
#my_labels =  c(">20%","20-80%",">80")

#my_probs = c(0 , 0.1, 0.2, 0.8, 0.9, 1)
#my_labels =  c(">10%", "10-20%", "20-80%", "80-90%", ">90%")

my_probs = c(0,0.025,0.2,0.8,0.975,1)
my_labels =  c("<2.5%","2.5-20%","20-80%","80-97.5%",">97.5%")

df <- df %>%
  mutate_at(vars(contains('SCORE')), list(SCALED = ~scale(.)[,1])) %>%
  rename_at(vars(contains('SCALED')), list(~str_remove(., "SCORE_"))) %>%
  mutate_at(vars(contains('SCORE')),
            list(CAT = ~cut(., breaks=quantile(., probs = my_probs),
                            labels=my_labels, include.lowest = T))) %>%
  rename_at(vars(contains('CAT')), list(~str_remove(., "SCORE_")))

#test <- df[1:20, 20:40]

```


## Properties of variables

Redefined name variables:

```{r}
names_oper <- list("REVASC", "CABG", "PCI")
names_ep <- list("AF", "INTRACRA", "MI", "STR", "GIBLEEDING")
```



**Sex and age**

```{r}

#Sex and age
df %>%
  summarise(N=n(),N_fem = sum(SEX_IMPUTED), fem_perc = mean(SEX_IMPUTED)*100, end_age = mean(FU_END_AGE), end_age_sd = sd(FU_END_AGE)) %>%
  knitr::kable() %>% kable_styling()

```


## Subset the data by operations

**Create datasets for each operations**

Lists with three datasets for each operations created

```{r}

#Datasets for each operation are filtered.
#As FinnGen assignes NA's to similar endpoints, we must replace NA's for
#AF and MI by zero, or they do not have observations.

dfs.oper <- lapply(names_oper, function(name){
    df %>% filter(get(name) == 1) %>%
    mutate(AF = if_else(is.na(AF), as.integer(0), AF),
           MI = if_else(is.na(MI), as.integer(0), MI))
  }) %>% setNames(names_oper) 


```

Sex and age

```{r}

#Sex and age

names_oper_age <- lapply(names_oper, function(name) {str_glue("{name}_AGE")})

mapply(
  function(dfi, name){
    dfi %>% summarise(N=n(), N_fem = sum(SEX_IMPUTED), fem_perc = mean(SEX_IMPUTED)*100, 
            mean_oper_age = mean(get(name)), sd_oper_age = sd(get(name)))},
  dfs.oper, names_oper_age) %>%
  as.matrix() %>% t() %>% as.data.frame() %>%
  mutate_at(.vars= c("fem_perc","mean_oper_age","sd_oper_age"), as.numeric) %>%
  knitr::kable(digits=2) %>%
  kable_styling()

```



## Subset data by operations and events

**Datasets for each combination of operation and PRS**

For each endpoint, *remove the prevalent cases* and create list of new
datasets. Both list of dataframes and long format dataframes are created for 
each operation.

```{r}

#Datasets for each operation are filtered and added to list

covs <- c("SEX_IMPUTED", "batch", "BL_YEAR","PC1", "PC2", "PC3", "PC4", "PC5", "PC6", "PC7", "PC8","PC9", "PC10")

#Lists of dataframes for each operations.
#PREVALENT ARE REMOVED

dfs <- lapply(names_oper, function(oper){
  my_create_sub_dfs(oper, dfs.oper, names_ep, covs)}) %>% 
  setNames(names_oper) 


```


## Characteristics tables for each operations

Here all characteristics are at the age of operation. 

```{r}

catVars <-  c("SEX_IMPUTED")
chr_rownames <- c("Age, years (SD)", "Women, n (%)")

chars <- 
  lapply(names_oper, function(oper){
    lapply(names_ep, function(ep){
      my_char_table(dfs[[oper]][[ep]], strata=unlist(ep), myVars=c(str_glue("{ep}_AGE"), "SEX_IMPUTED"), catVars, chr_rownames, name="N") %>%
        rbind(Endpoint=c("","",""), .) %>%
        set_rownames(., c(ep, "N", chr_rownames)) %>%
        set_colnames(., c("Overall", "Cases", "Controls"))
    }) %>% setNames(names_ep)
  }) %>% setNames(names_oper)

#REVASC
do.call(rbind, chars$REVASC) %>%
  knitr::kable() %>% kable_styling()

#CABG
do.call(rbind, chars$CABG) %>%
  knitr::kable() %>% kable_styling()

#PCI
do.call(rbind, chars$PCI) %>%
  knitr::kable() %>% kable_styling()

```




# Survival analysis

**For each three operation and six endpoints: **

* Take patients after the operation.
* Remove prevalent cases of the endpoint
* Survival analysis:
  + Hazard ratios per 1 SD change: table & curves
* Covariates: SEX_IMPUTED, batch, BL_YEAR, PC1:PC10

## Proportional hazard assumption: log log curve

<details><summary>**log log curve**</summary>

**Model by survfit**

```{r}

kms <- 
  lapply(names_oper, function(oper){
    lapply(names_ep, function(ep){
      survfit(Surv(AGE_DIFF, get(ep)) ~ get(str_glue("{ep}_CAT")), data=dfs[[oper]][[ep]])
    }) %>% setNames(names_ep)
  }) %>% setNames(names_oper) 

```


**KM plots**

```{r, fig.width=10, fig.height=5, eval=F}

for(oper in unlist(names_oper)) {
  ep=names_ep[[1]]   #Tämän alustuksen tarve on kyllä ihme juttu. Suoritus kaatuu ilman sitä.
  
  plot_list <- lapply(names_ep, function(ep){
    print(str_glue("{oper}-{ep}"))
    my_ggkmplot(kms[[oper]][[ep]], str_glue("{oper}-{ep}"), my_labels, isFive=T) 
  }) 
  plots <- arrange_ggsurvplots(plot_list, nrow = 2, ncol = 3, title = "Kaplan Meier curves")
  ggsave(file = str_glue("{fig_path}/km.prs.after.{oper}.png"), plot = plots, height = 6, width = 12, dpi = 200)
}


```



**log log curve**


```{r, fig.width=10, fig.height=8}

for(oper in unlist(names_oper)) {
  ep=names_ep[[1]]   #Tämän alustuksen tarve on kyllä ihme juttu. Suoritus kaatuu ilman sitä.
  
  plot_list <- lapply(names_ep, function(ep){
    print(str_glue("{oper}-{ep}"))
    my_ggkmplot(kms[[oper]][[ep]], str_glue("{oper}-{ep}"), my_labels, fun="cloglog", xlim=c(0.1,30), isFive=T)
    
  }) 
  plots <- arrange_ggsurvplots(plot_list, nrow = 2, ncol = 3, title = "log-log curves")
  ggsave(file = str_glue("{fig_path}/log.prs.after.{oper}.png"), plot = plots, height = 6, width = 12, dpi = 200)
}


```

Not very convincing. AF seems OK. Perhaps also others roughly parallel, but not really different. 

</details>
<br>


## Cox model: covariate adjusted, continuous score


**Running models**


```{r}
covs_f<- paste(covs, collapse = " + ")

cxs.cs <- 
  lapply(names_oper, function(oper){
    lapply(names_ep, function(ep){
      my_formula <- str_glue("Surv(AGE_DIFF, {ep}) ~ {ep}_SCALED + {oper}_AGE + {covs_f}")
      coxph(as.formula(my_formula), data=dfs[[oper]][[ep]])
    }) %>% setNames(names_ep)
  }) %>% setNames(names_oper) 

```


**Table: Hazard rations**

*Main results*

```{r}

oper="REVASC"
my_cxlist_to_hrtable(cxs.cs[[oper]], oper) %>% 
    knitr::kable(digits=2) %>% kable_styling()

oper="CABG"
my_cxlist_to_hrtable(cxs.cs[[oper]], oper) %>% 
    knitr::kable(digits=2) %>% kable_styling()

oper="PCI"
my_cxlist_to_hrtable(cxs.cs[[oper]], oper) %>% 
    knitr::kable(digits=2) %>% kable_styling()

```

After all operations, AF PRS shows significant associations to the survival time before first AF event.
After bonferroni correction (0.05/15=0.003) other endpoints are not statistically significant (except CABG-STR and CABG-MI).  


*Sex & age of operation included(REVASC only)*

```{r}

#SEX and AGE for REVASC
my_cxlist_to_hrtable_selected(cxs.cs$REVASC, "REVASC", select="SEX|AGE") %>% 
  knitr::kable(digits=2) %>%
    kable_styling()

```

Operation age is significant in all cases. Sex is significant only in case of AF.



<details><summary>**All coefficients**</summary>


```{r, eval=T}

lapply(cxs.cs, print)

```

Genetic covariates are not significant. Only very few exceptions. 


</details>
<br>




**Cox-plot**

Creating covariate combinations for the plot.


```{r}

var_qs= c(0.01, 0.1, 0.5, 0.9, 0.99)  #Vastaa suunnilleen niitä meidän kategorioita...
#var_qs=c(0.1, 0.2, 0.5, 0.8, 0.9)

exp.cs.cxs <- 
  lapply(names_oper, function(oper){
    lapply(names_ep, function(ep){
      ep_scaled <- str_glue("{ep}_SCALED") 
      my_expand_cs(cxs.cs[[oper]][[ep]], ep_scaled, dfs[[oper]][[ep]], 
                   batch = "AxiomGT1_b44_V2.calls", agename=str_glue("{oper}_AGE"), 
                   var_qs = quantile(dfs[[oper]][[ep]][[ep_scaled]], var_qs))
    }) %>% setNames(names_ep)
  }) %>% setNames(names_oper) 


fit.cs.cxs <- 
  lapply(names_oper, function(oper){
    lapply(names_ep, function(ep){
        survfit(cxs.cs[[oper]][[ep]], newdata = exp.cs.cxs[[oper]][[ep]])
  }) %>% setNames(names_ep)
}) %>% setNames(names_oper) 
    

```

And the plot:

```{r, fig.width=10, fig.height=5}

my_labels = c("1%", "10%", "50%", "90%", "99%")

for(oper in unlist(names_oper)) {
  ep=names_ep[[1]]   #Tämän alustuksen tarve on kyllä ihme juttu. Suoritus kaatuu ilman sitä.
  
  plot_list <- lapply(names_ep, function(ep){
    #print(str_glue("{oper}-{ep}"))
    my_ggcoxplot(fit.cs.cxs[[oper]][[ep]], exp.cs.cxs[[oper]][[ep]], str_glue("{oper}-{ep}"), my_labels, xlim=c(0,30), isFive=T)
  }) 
  plots <- arrange_ggsurvplots(plot_list, nrow = 2, ncol = 3, title = "Cox-plots")
  ggsave(file = str_glue("{fig_path}/cx.cs.prs.after.{oper}.png"), plot = plots, height = 6, width = 12, dpi = 200)
}
```



<details><summary>**Proportional hazard assumption**</summary>


Schoenfeld residuals against the transformed time


```{r}
extract_ph_test <- function(cx){
  tmp <- cox.zph(cx)
  tmp$table[c(1,16),]
}

lapply(names_oper, function(oper){
  lapply(names_ep, function(ep){
    tryCatch(extract_ph_test(cxs.cs[[oper]][[ep]]), error=function(e) NULL)
  }) %>% setNames(names_ep)
}) %>% setNames(names_oper) 
  

lapply(names_ep, function(ep){
    tryCatch(cox.zph(cxs.cs[["REVASC"]][[ep]]), error=function(e) NULL)
}) %>% setNames(names_ep)


```


</details>
<br>
