---
title: "AF PRS vs. AF after revascularization"
editor_options:
  chunk_output_type: console
output:
  html_document:
    number_sections: true
---

```{r setup, include=FALSE}
#packrat::init()
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```


**Libraries**


```{r}
#setwd("/finngen/red/cabg_prs")

#install.packages("/home/ivm/R/src/packrat_0.5.0.tar.gz", repos=NULL, type="source")
#install.packages("/finngen/green/tableone_0.12.0.tar.gz", repos=NULL, type="source")
#install.packages("/finngen/green/compareC_1.3.1.tar.gz", repos=NULL, type="source")

#packrat::on()
library(tidyverse)
library(survival)   # survival analysis

library(data.table) # fread() function
library(gridExtra)  # plots to grid
library(survminer)  # ggbased visualization and extra diagnostics
#library(survRM2)    # restricted mean survival times
#library(visdat)     # visualization of tibble and na's
##library(forestplot) # forestplot
library(compareC)   # compare c-index
library(kableExtra) # Pretty tables
library(tableone)   # Characteristics table

source('/finngen/red/cabg_prs/scripts/functions2.R')

data_path <- "/finngen/red/cabg_prs/data"
prs_path <- "/finngen/red/cabg_prs/data/ukb_prs"
fig_path <- "/finngen/red/cabg_prs/figs"

```


# Data

## Variables

**Operations**

* I9_CABG, Coronary artery bypass crafting
* I9_ANGIO, Coronary angiopasty
* I9_REVASC, I9_CABG OR I9_ANGIO

Three samples: patients *after* CABG, ANGIO and REVASC.

**For polygenic riskscores**

* I9_AF (ukb: 20002_1471), Atrial fibrillation

**For clinical riskscore CHA~2~DS~2~-VASc:**

* Congestive heart failure = I9_HEARTFAIL
* Hypertension = I9_HYPTENS
* Diabetes mellitus = E4_DIABETES
* Stroke, TIA, or TE = C_STROKE
* Vascular disease = I9_MI or I9_PAD

## Download & preprocess data


**Copying and preprocessing outside R**

Endpoint and covariate files transfered and unzipped. Columns selected from endpoint file prior to import to r, because the original phenotype file is very large.


```{bash, eval=F}

#Endpoint file and covariate files unzipped
phenodir="/finngen/red/cabg_prs/data/phenotypes"
gzip -dc /finngen/library-red/finngen_R6/phenotype_2.0/data/finngen_R6_v2_endpoint.gz > $phenodir/finngen_R6_v2_endpoint
gzip -dc /finngen/library-red/finngen_R6/phenotype_2.0/data/finngen_R6_cov_pheno_1.0.txt.gz > $phenodir/finngen_R6_cov_pheno_1.0.txt

#Self written perl-script to extract columns from phenotype file (runned from directory 'scripts')
phenodir="/finngen/red/cabg_prs/data/phenotypes"
perl select_columns.pl $phenodir/finngen_R6_v2_endpoint        $phenodir/finngen_R6_v2_CABG    fg_pheno_cols.txt &
perl select_columns.pl $phenodir/finngen_R6_cov_pheno_1.0.txt  $phenodir/R6_cov_pheno_selected fg_cov_cols.txt &
perl select_columns.pl $phenodir/finngen_R6_v2_endpoint       $phenodir/finngen_R6_v2_chadsvasc    fg_chadsvasc_cols.txt &
#perl select_columns.pl $phenodir/finngen_R6_cov_pheno_1.0.txt  $phenodir/finngen_R6_v2_chadsvasc    fg_chadsvasc_cols.txt &

```


**Loading and combining the data**

Downloading the data.

```{r}

names_in <- list("I9_AF")

#PRS scores
prs_df <- lapply(names_in, function(name) {
  fread(str_glue("{prs_path}/ukb.{name}.both_sexes.profile")) %>%
        select(IID, SCORE) %>%
      rename(FINNGENID = IID, !!paste(name, "SCORE", sep="_") := SCORE)}
  ) %>%
  reduce(left_join, by = "FINNGENID")

#endpoints and covariates
phenotypes <- fread(str_glue("{data_path}/phenotypes/finngen_R6_v2_CABG"), sep = "\t") 
covs_in <- fread(str_glue("{data_path}/phenotypes/R6_cov_pheno_selected"), sep = "\t")
pheno_chdvsc <- fread(str_glue("{data_path}/phenotypes/finngen_R6_v2_chadsvasc"), sep = "\t") %>%
  select(-contains('MI')) #To avoid duplicates

```

Combining the data.

```{r}
df <- phenotypes %>%
  left_join(prs_df, by = "FINNGENID") %>%
  left_join(covs_in, by = "FINNGENID") %>%
  left_join(pheno_chdvsc, by = "FINNGENID") %>%
  as_tibble() %>%
  mutate_at(c("batch","cohort"),as.factor) 

```


**Preprocess**

Removing individuals without genetic information.
Also endpointnames are simplified.

```{r}
#
#glimpse(df)
#vis_dat(sample_n(df, 10000))
df <- df %>% filter(!is.na(batch)) %>%
  rename_at(vars(contains('I9_')), list(~str_remove(., "I9_"))) %>%
  rename_at(vars(contains('K11_')), list(~str_remove(., "K11_"))) %>%
  rename_at(vars(contains('E4_')), list(~str_remove(., "E4_"))) %>%
  rename_at(vars(contains('C_STRO')), list(~str_remove(., "C_"))) %>%
  rename_at(vars(contains('ANGIO')), list(~str_replace(., "ANGIO", "PCI")))
#vis_dat(sample_n(df, 10000))

#Number of individuals:
#dim(df)[1]

```


Creating scaled & categorical variables.


```{r}

#my_probs = c(0 ,0.2, 0.8, 1)
#my_labels =  c(">20%","20-80%",">80")

#my_probs = c(0 , 0.1, 0.2, 0.8, 0.9, 1)
#my_labels =  c(">10%", "10-20%", "20-80%", "80-90%", ">90%")

my_probs = c(0,0.025,0.2,0.8,0.975,1)
my_labels =  c("<2.5%","2.5-20%","20-80%","80-97.5%",">97.5%")

df <- df %>%
  mutate(AF_SCALED = scale(AF_SCORE),
         AF_CAT = cut(AF_SCORE, breaks=quantile(df$AF_SCORE, my_probs),
                            labels=my_labels, include.lowest = T))

```


**Variable names redefined**


```{r}
names_oper <- list("REVASC", "CABG", "PCI")
names_ep <- list("AF")
```


**Number of individuals, AF, sex and age**

Age is age at end of followup.

```{r}

#Sex and age
df %>%
  summarise(N=n(), N_AF = sum(AF, na.rm = T), AF_perc = mean(AF, na.rm = T)*100 ,N_fem = sum(SEX_IMPUTED), fem_perc = mean(SEX_IMPUTED)*100, end_age = mean(FU_END_AGE), end_age_sd = sd(FU_END_AGE)) %>%
    knitr::kable(digits=1) %>%  kable_styling()

```


## Subset the data by operations

**Create datasets for each operations**

Lists with three datasets for each operations created

```{r}

#Datasets for each operation are filtered.
#As FinnGen assignes NA's to similar endpoints, we must replace NA's for
#AF and MI by zero, or they do not have observations.

dfs.oper <- lapply(names_oper, function(name){
    df %>% filter(get(name) == 1) %>%
    mutate(AF = if_else(is.na(AF), as.integer(0), AF),
           MI = if_else(is.na(MI), as.integer(0), MI))
  }) %>% setNames(names_oper) 


```


**CHA~2~DS~2~-VASc score**

```{r}

#Chadsvasc score:

dfs.oper <- lapply(names_oper, function(oper){
  
  oper_age <- str_glue("{oper}_AGE")
  dfs.oper[[oper]] %>%
  mutate(MIPAD = if_else(MI|PAD, 1, 0)) %>%
  rowwise()%>%
  mutate(MIPAD_AGE = min(MI_AGE, PAD_AGE)) %>%
  ungroup() %>%
  mutate(CVSC_AGE_65 = if_else(get(oper_age) >= 65, 1, 0),   
         CVSC_AGE_75 = if_else(get(oper_age) >= 75, 1, 0),
         CVSC_AGE_65_75 = if_else((CVSC_AGE_65 == 1 & CVSC_AGE_75 == 0),1, 0),
         CVSC_HEARTFAIL = if_else(HEARTFAIL_AGE < get(oper_age), 1, 0),
         CVSC_HYPTENS = if_else(HYPTENS_AGE < get(oper_age), 1, 0),
         CVSC_DIABETES = if_else(DIABETES_AGE < get(oper_age), 1, 0),
         CVSC_STROKE = if_else(STROKE_AGE < get(oper_age), 1, 0),
         CVSC_MIPAD = if_else(MIPAD_AGE < get(oper_age), 1, 0), 
         preop_VHD = if_else(VHD_AGE < get(oper_age), 1, 0),                      
         preop_OBESITY = if_else(OBESITY_AGE < get(oper_age), 1, 0),
         CVSC = CVSC_HEARTFAIL + CVSC_HYPTENS + CVSC_AGE_75 + CVSC_DIABETES + 
           2*CVSC_STROKE + CVSC_MIPAD + CVSC_AGE_65 + SEX_IMPUTED, 
         CVSC_SCALED = scale(CVSC))
  
}) %>% setNames(names_oper) 

```

Correlation between clinical and genetic score:

```{r}

#Pearson
lapply(names_oper, function(oper){
  cor(dfs.oper[[oper]][["CVSC_SCALED"]],dfs.oper[[oper]][["AF_SCALED"]], method = "pearson")
})%>% setNames(names_oper) %>% unlist()

#Spearman
lapply(names_oper, function(oper){
  cor(dfs.oper[[oper]][["CVSC_SCALED"]],dfs.oper[[oper]][["AF_SCALED"]], method = "spearman")
})%>% setNames(names_oper) %>% unlist()

```


**Prevalent cases removed**

```{r}

#Datasets for each operation are filtered and added to list

covs <- c("SEX_IMPUTED", "batch", "BL_YEAR","PC1", "PC2", "PC3", "PC4", "PC5", "PC6", "PC7", "PC8","PC9", "PC10")

#Lists of dataframes for each operations and each endpoint of interest
#PREVALENT ARE REMOVED

dfs <- lapply(names_oper, function(oper){
  my_create_sub_dfs(oper, dfs.oper, names_ep, covs) }) %>% 
  setNames(names_oper) 

#list of dataframes is simplified as we have now only one endpoint of interest, AF
dfs.af <- lapply(names_oper, function(oper){
  dfs[[oper]][["AF"]] }) %>%
  setNames(names_oper) 


```

**Mean followup time**

```{r}

lapply(names_oper, function(oper){
  dfs.af[[oper]] %>%
    summarise(mean = mean(AGE_DIFF), median = median(AGE_DIFF))
})%>% setNames(names_oper)


```




## Characteristics tables for each operations

Here all characteristics are at the age of operation. 

```{r}

myVars <-   c("AGE_DIFF", "AF_AGE", "SEX_IMPUTED", "CVSC", "CVSC_HEARTFAIL", "CVSC_HYPTENS", "CVSC_AGE_75", "CVSC_DIABETES", 
              "CVSC_STROKE", "CVSC_MIPAD", "CVSC_AGE_65_75", "preop_VHD", "preop_OBESITY")

catVars <-  c("SEX_IMPUTED", "CVSC_HEARTFAIL", "CVSC_HYPTENS", "CVSC_AGE_75", "CVSC_DIABETES", 
              "CVSC_STROKE", "CVSC_MIPAD", "CVSC_AGE_65_75", "preop_VHD", "preop_OBESITY")

chr_rownames <- c("Follow up, years (SD)", "Age, years (SD)", "Women, n (%)", "CHA2DS2-VASc score (SD)", "Hearth failure, n (%)", "Hypertension, n (%)", "Age > 74, n (%)",
              "Diabetes, n (%)", "Prior stroke or TIA, n (%)", "Vascular disease, n (%)", "Age 65-75, n (%)")

my_char_table(dfs.af[["REVASC"]], "AF", myVars, catVars, chr_rownames, name="REVASC") %>%
  knitr::kable() %>% kable_styling()

my_char_table(dfs.af[["CABG"]], "AF", myVars, catVars, chr_rownames, name="CABG") %>%
  knitr::kable() %>% kable_styling()

my_char_table(dfs.af[["PCI"]], "AF", myVars, catVars, chr_rownames, name="PCI") %>%
  knitr::kable() %>% kable_styling()

```

## Linearized rate

```{r}

dfs.af[["REVASC"]] %>%
    group_by(AF_CAT) %>%
    summarise((sum(AF)/sum(FU_END_AGE-REVASC_AGE))*100)
  
dfs.af[["CABG"]] %>%
    group_by(AF_CAT) %>%
    summarise((sum(AF)/sum(FU_END_AGE-CABG_AGE))*100)

dfs.af[["PCI"]] %>%
    group_by(AF_CAT) %>%
    summarise((sum(AF)/sum(FU_END_AGE-PCI_AGE))*100)



```




# Survival analysis

**AF after each three operation : **

* Model1: Covariates: {oper}_AGE, SEX_IMPUTED, batch, BL_YEAR, PC1:PC10
* Model2: Covariates: CHA~2~DS~2~-VASc, {oper}_AGE, SEX_IMPUTED, batch, BL_YEAR, PC1:PC10
  

## Proportional hazard assumption: log log curve

<details><summary>**KM & log log curves**</summary>

**Model by survfit**

```{r}

kms <- 
  lapply(names_oper, function(oper){
    lapply(names_ep, function(ep){
      survfit(Surv(AGE_DIFF, get(ep)) ~ get(str_glue("{ep}_CAT")), data=dfs[[oper]][[ep]])
    }) %>% setNames(names_ep)
  }) %>% setNames(names_oper) 

```


**KM plots**

```{r, fig.width=5, fig.height=4}

for(oper in unlist(names_oper)) {
  ep=names_ep[[1]]   #Tämän alustuksen tarve on kyllä ihme juttu. Suoritus kaatuu ilman sitä.
  
  plot_list <- lapply(names_ep, function(ep){
    #print(str_glue("{oper}-{ep}"))
    my_ggkmplot(kms[[oper]][[ep]], str_glue("{oper}-{ep}"), my_labels, isFive=T) 
  }) 
  plots <- arrange_ggsurvplots(plot_list, nrow = 1, ncol = 1, title = "Kaplan Meier curves")
  #ggsave(file = str_glue("{fig_path}/km.prs.after.{oper}.png"), plot = plots, height = 6, width = 12, dpi = 200)
}


```



**log log curve**


```{r, fig.width=5, fig.height=4}

for(oper in unlist(names_oper)) {
  ep=names_ep[[1]]   #Tämän alustuksen tarve on kyllä ihme juttu. Suoritus kaatuu ilman sitä.
  
  plot_list <- lapply(names_ep, function(ep){
    #print(str_glue("{oper}-{ep}"))
    my_ggkmplot(kms[[oper]][[ep]], str_glue("{oper}-{ep}"), my_labels, fun="cloglog", xlim=c(0.1,30), isFive=T)
    
  }) 
  plots <- arrange_ggsurvplots(plot_list, nrow = 1, ncol = 1, title = "log-log curves")
  #ggsave(file = str_glue("{fig_path}/log.prs.after.{oper}.png"), plot = plots, height = 6, width = 12, dpi = 200)
}


```


</details>
<br>


## Cox model: covariate adjusted, continuous score


**Running models**


```{r}
covs_f<- paste(covs, collapse = " + ")

# Only PRS
ep <- "AF"

cxs.cs.both <- 
  lapply(names_oper, function(oper){
      my_formula <- str_glue("Surv(AGE_DIFF, {ep}) ~ {ep}_SCALED + CVSC_SCALED + {oper}_AGE + {covs_f}")
      coxph(as.formula(my_formula), data=dfs.af[[oper]])
  }) %>% setNames(names_oper) 


```


**Table: Hazard rations**


```{r}

hr.cs.tables <- 
  lapply(names_oper, function(oper){

  nN <- dfs.af[[oper]] %>% 
    summarize(sum(AF), n()-sum(AF)) %>%
    unite("nN", sep="/")
  
  my_extr_coef(cxs.cs.both[[oper]], "PRS", select="AF_SCALED") %>%
    my_tidy_table() %>%
    cbind(.,nN) %>%
    select(name, est, pval, nN) %>%
    rename(!!oper := name, `HR (95% CI)` = est, `P-value` = pval, `Cases/Controls` = nN) #%>%
    #mutate("P-value" = sub("e(-\\d+)", "x10<sup>\\1</sup>", `P-value`, perl=T))   #html-version/proper exponents
  
}) %>% setNames(names_oper)

hr.cs.tables$REVASC %>% knitr::kable() %>% kable_styling()
hr.cs.tables$CABG %>% knitr::kable() %>% kable_styling()
hr.cs.tables$PCI %>% knitr::kable() %>% kable_styling()

```


*Sex & age of operation included*

```{r}

#SEX and AGE
my_cxlist_to_hrtable_selected(cxs.cs.both, "name", select="SEX|AGE") %>% 
  knitr::kable(digits=2) %>%
  kable_styling()

```



<details><summary>**All coefficients**</summary>


```{r}

#lapply(cxs.cs.prs, print)
lapply(cxs.cs.both, print)
#lapply(cxs.cs.cvsc, print)

```

Genetic covariates are not significant. Only very few exceptions. 


</details>
<br>

<details><summary>**Proportional hazard assumption**</summary>

Schoenfeld residuals against the transformed time


```{r}
extract_ph_test <- function(cx){
  tmp <- cox.zph(cx)
  tmp$table[c(1,2, 3, 16),]
}

lapply(names_oper, function(oper){
  tryCatch(extract_ph_test(cxs.cs.both[[oper]]), error=function(e) NULL)
}) %>% setNames(names_oper) 

```

</details>
<br>


## Cox model: covariate adjusted, categorized score


**Running models**


```{r}

ep="AF"

cxs.both <- 
  lapply(names_oper, function(oper){
      df_tmp <- dfs.af[[oper]] %>%
        mutate(!!str_glue("{ep}_CAT") := relevel(get(str_glue("{ep}_CAT")), ref = "20-80%"))
      my_formula <- str_glue("Surv(AGE_DIFF, {ep}) ~ {ep}_CAT + CVSC_SCALED + {oper}_AGE + {covs_f}")
      coxph(as.formula(my_formula), data=df_tmp)
  }) %>% setNames(names_oper) 

```


**Table: Hazard ratios:**


```{r, fig.width=12}

#Model 2
hr2 <- lapply(names_oper, function(oper){
  my_cx_to_hrtable_cat(cxs.both[[oper]], dfs.af[[oper]], ep, select="_CAT", title=oper)
})  %>% setNames(names_oper) 
#  do.call(cbind, .) 

#Combined. Here repeated extra columns with quantilies are removed
hr2$REVASC %>%  knitr::kable() %>% kable_styling()
hr2$CABG %>%  knitr::kable() %>% kable_styling()
hr2$PCI %>%  knitr::kable() %>% kable_styling()

```


**Selecting covariates for plot**

We used **mean** values for all numeric variables. For 'batch', we use the level with **lowest significance**.


<details><summary>**Choosing covariate values**</summary>


The batches were not significant, so this does not really matter. 
Let us take lowest significance from a one example case: REVASC-AF

```{r}

cx.test <- cxs.both$REVASC

summary(cx.test)$coefficients %>%
  as.data.frame() %>%
  mutate(name = row.names(.)) %>%
  rename(pval="Pr(>|z|)") %>%
  select("name", "coef", "exp(coef)","se(coef)", "pval") %>%  
  arrange(pval)



```

Batch: AxiomGT1_b44_V2.calls had both low p-value and effect size. It is chosen.

</details>
<br>


**Cox-plot**

Creating covariate combinations for the plot.


```{r}

exp.cxs2 <- 
  lapply(names_oper, function(oper){

    exp.tmp.covs <-  my_expand_covs(dfs.af[[oper]], batch = "AxiomGT1_b44_V2.calls")
    exp.tmp.vars <- expand.grid(AF_CAT = levels(dfs.af[[oper]][["AF_CAT"]]),
                                CVSC_SCALED = mean(dfs.af[[oper]][["CVSC_SCALED"]]),
                                TEST = mean(dfs.af[[oper]][[str_glue("{oper}_AGE")]]),
                                SEX_IMPUTED = 0.5) %>%
                    rename(!!str_glue("{oper}_AGE") := TEST)
   
    bind_cols(exp.tmp.vars, exp.tmp.covs)
    
  }) %>% setNames(names_oper)   


fit.cxs2 <- 
  lapply(names_oper, function(oper){
        survfit(cxs.both[[oper]], newdata = exp.cxs2[[oper]])
}) %>% setNames(names_oper) 

```


And the plot:

```{r, fig.width=12, fig.height=4}

oper="REVASC"
p1 <- my_ggcoxplot(fit.cxs2[[oper]], exp.cxs2[[oper]], "Revascularization", rev(my_labels), xlim=c(0,30), ylim=c(0,0.9), legend = "none")[["plot"]]

oper="CABG"
p2 <- my_ggcoxplot(fit.cxs2[[oper]], exp.cxs2[[oper]], str_glue("{oper}"), my_labels, xlim=c(0,30), ylim=c(0,0.9), legend="none", ylab="")[["plot"]]

oper="PCI"
p3 <- my_ggcoxplot(fit.cxs2[[oper]], exp.cxs2[[oper]], str_glue("{oper}"), my_labels, xlim=c(0,30), ylim=c(0,0.9), legend="none", ylab="")[["plot"]] 

lgnd <- my_ggcoxplot(fit.cxs2[[oper]], exp.cxs2[[oper]], "Revascularization", rev(my_labels), legend="right", legend.title="AF PRS")[["plot"]] %>%
         get_legend()

plots2 <- grid.arrange(p1, p2, p3, lgnd, ncol = 4, widths = c(3,3,3,1))

ggsave(file = str_glue("{fig_path}/cx.prs.af2.png"), plot = plots2, height = 3, width = 10, dpi = 300)
ggsave(file = str_glue("{fig_path}/cx.prs.af2.eps"), plot = plots2, device = cairo_ps, height = 3, width = 10, dpi = 300)


```




## C-statistics for AF



## Cox model: covariate adjusted, continuous score


**Running models**


```{r}
covs_f<- paste(covs, collapse = " + ")

# Only PRS
ep <- "AF"


# Model1, only CHADSVASC
cxs.cs.m1 <- 
  lapply(names_oper, function(oper){
      my_formula <- str_glue("Surv(AGE_DIFF, {ep}) ~ CVSC_SCALED + {oper}_AGE + {covs_f}")
      coxph(as.formula(my_formula), data=dfs.af[[oper]])
  }) %>% setNames(names_oper) 



# Model2, CHADSVASC + VHD + OBESITY
cxs.cs.m2 <- 
  lapply(names_oper, function(oper){
      my_formula <- str_glue("Surv(AGE_DIFF, {ep}) ~ CVSC_SCALED + preop_OBESITY + preop_VHD + {oper}_AGE + {covs_f}")
      coxph(as.formula(my_formula), data=dfs.af[[oper]])
  }) %>% setNames(names_oper) 



# Model3, PRS + CHADSVASC + VHD + OBESITY
cxs.cs.m3 <- 
  lapply(names_oper, function(oper){
      my_formula <- str_glue("Surv(AGE_DIFF, {ep}) ~ {ep}_SCALED + CVSC_SCALED + preop_OBESITY + preop_VHD + {oper}_AGE + {covs_f}")
      coxph(as.formula(my_formula), data=dfs.af[[oper]])
  }) %>% setNames(names_oper) 

```

**HR/1SD change:**


CA~2~HDS~2~-VASc only:

```{r}
#Model1
my_cxlist_to_hrtable_selected(cxs.cs.m1, "Operation", select="_SCALED|preop") %>% 
  knitr::kable(digits=2) %>%
  kable_styling()

```



CA~2~HDS~2~-VASc + obesity + VHD:

```{r}

my_cxlist_to_hrtable_selected(cxs.cs.m2, "Operation", select="_SCALED|preop") %>% 
  knitr::kable(digits=2) %>%
  kable_styling()

```


CA~2~HDS~2~-VASc + obesity + VHD + PRS:

```{r}

my_cxlist_to_hrtable_selected(cxs.cs.m3, "Operation", select="_SCALED|preop") %>% 
  knitr::kable(digits=2) %>%
  kable_styling()

```

**Cox plot: CA~2~HDS~2~-VASc + PRS**

Creating covariate combinations for the plot.

```{r}

exp.cs.cxs3 <- 
  lapply(names_oper, function(oper){
       my_expand_cs_cvsc3(cxs.m3[[oper]],  "AF_SCALED", dfs.af[[oper]], 
                  batch = "AxiomGT1_b44_V2.calls", agename=str_glue("{oper}_AGE"), 
                  #var_qs = quantile(dfs.af[[oper]][["AF_SCALED"]], c(0.8413, 0.1587)), 
                  #scr_qs = quantile(dfs.af[[oper]][["CVSC_SCALED"]], c(0.8413, 0.1587)))
                  var_qs = quantile(dfs.af[[oper]][["AF_SCALED"]], c(0.9, 0.1)), 
                  scr_qs = quantile(dfs.af[[oper]][["CVSC_SCALED"]], c(0.9, 0.1)))
  }) %>% setNames(names_oper) 


fit.cs.cxs3 <- 
  lapply(names_oper, function(oper){
        survfit(cxs.cs.m3[[oper]], newdata = exp.cs.cxs3[[oper]])
}) %>% setNames(names_oper) 


```



And the plot:

```{r, fig.width=12, fig.height=4}

#label_name = expression('AF PRS - CHA'[2]*'DS'[2]*'-VASc')
label_name = expression('AF PRS/CHA'[2]*'DS'[2]*'-VASc')
#lab_tmp = c("90% - 90%", "10% - 90%", "90% - 10%", "10% - 10%")
lab_tmp = c("1SD/1SD", "-1SD/1SD", "1SD/-1SD", "-1SD/-1SD")
#palette=c("#C77CFF",  "#00BFC4",  "#7CAE00", "#F8766D") #ggplot default reversed

oper="REVASC"
p1 <- ggsurvplot(fit.cs.cxs3[[oper]], data =  exp.cs.cxs3[[oper]], 
               fun = "event", censor = F, conf.int=F, ggtheme = theme_bw(),
               xlim=c(0,30), ylim=c(0,0.9), break.time.by=5, legend="none",
               size=0.5, font.tickslab=9.5, font.legend=9.5, title="Revascularization")[["plot"]]

oper="CABG"
p2 <- ggsurvplot(fit.cs.cxs3[[oper]], data =  exp.cs.cxs3[[oper]], 
               fun = "event", censor = F, conf.int=F, ggtheme = theme_bw(),
               xlim=c(0,30), ylim=c(0,0.9), break.time.by=5, legend="none",
               size=0.5, font.tickslab=9.5, font.legend=9.5, title=oper)[["plot"]]

oper="PCI"
p3 <- ggsurvplot(fit.cs.cxs3[[oper]], data =  exp.cs.cxs3[[oper]], 
               fun = "event", censor = F, conf.int=F, ggtheme = theme_bw(),
               xlim=c(0,30), ylim=c(0,0.9), break.time.by=5, legend="none",
               size=0.5, font.tickslab=9.5, font.legend=9.5, title=oper)[["plot"]]

lgnd <- ggsurvplot(fit.cs.cxs3[[oper]], data =  exp.cs.cxs3[[oper]], censor = F,
                   legend.title=label_name, legend.labs=lab_tmp, legend="right",
                   size=0.5, font.tickslab=9.5, font.legend=9.5)[["plot"]] %>%
        get_legend()

plots2 <- grid.arrange(p1,p2, p3, lgnd, ncol = 4, widths = c(3,3,3,2))

#ggsave(file = str_glue("{fig_path}/cx.prs.af.m3sd.png"), plot = plots2, height = 3, width = 11, dpi = 300)
#ggsave(file = str_glue("{fig_path}/cx.prs.af.m3sd.eps"), plot = plots2, device = cairo_ps, height = 3, width = 10.6, dpi = 300)


```



**C-index**

```{r}

cindex_table <- 
  lapply(names_oper, function(oper){
    compc<- compareC(timeX = dfs.af[[oper]][["AGE_DIFF"]],
                     statusX = dfs.af[[oper]][["AF"]],
                     scoreY= 1 - cxs.cs.m3[[oper]][["linear.predictors"]], # '-' kääntää järjestyksen
                     scoreZ= 1 - cxs.cs.m2[[oper]][["linear.predictors"]])
    est <- prettify_estimate(estimate=compc$est.diff, stderr=sqrt(compc$est.vardiff_c), r=3)
    tibble(Name=oper, m1=round(cxs.cs.m1[[oper]]$concordance[6],4), Cxy=round(compc$est.c[1],4), Cxz=round(compc$est.c[2],4), 
           diff_est=est, diff_pval=signif(compc$pval,1))
}) %>%
do.call(rbind, .) %>%
select(Name, m1, Cxz, Cxy, diff_est, diff_pval) %>%
rename(Procedure=Name, `Model 1`=m1, `Model 3`=Cxy, `Model 2`=Cxz, `Difference (95% CI)`=diff_est, `P-value`=diff_pval) 


cindex_table %>%
knitr::kable() %>%
  kable_styling()

#Felix: En nyt ole varma miksi tuo standard error lasketaan ottamalla varianssista neliöjuuri, eikä myös jakamalla samplen koon neliöjuurella. 'Normaalistihan' tuossa kaavassa pitäisi olla jakajana sqrt(var)/sqrt(n) eikä pelkkä sqrt(var). Koska CI myötäilee P-arvoa tapauksessa P=0.05, niin olen aika varma.


```

.
